---
title: "Non-parametric GaPP"
output: bookdown::html_document2
link-citations: yes
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{gamrel-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: nonparametric-bib.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gamrel)
```

# Background

A first attempt was written up in @Arnold_Chukova_Hayakawa_APARM:2020.

## Errata: 

  * In eq. (17) $\alpha_1$ should be $\alpha$ [R code was correct]
  * In eq. (19) the exponent $a_1+K$ should be $a_1+K-1$ [R code was incorrect, but this effectively changed the value of $a_1$ in the hyperprior specification]
  * In eq. (19) the expression $\log w_K$ should be $\log \tilde{w}_K$ [R code was correct]
  * In the description of the simulation process in \Sec IV the locations (and their associated weigts) need to be sorted so that the locations $\theta_k$ are in ascending order [R code was correct]

# Model

## Basic Model Specification

Times to failure are drawn from an inhomogenous hazard rate function $\lambda(t)$.  Failures are iid, with density $f(t|\lambda)$ and survival function $\bar{F}(t|\lambda)$
\begin{eqnarray}
   f(t | \lambda) &=& \lambda(t) e^{-\Lambda(t)}    
   (\#eq:density)\\
   \bar{F}(t | \lambda) &=& e^{-\Lambda(t)}
   (\#eq:survival-function)
\end{eqnarray}
where
\begin{equation}
   \Lambda(t) = \int_0^t \lambda(u)\,{\rm d}u  
   (\#eq:cumulative-hazard)
\end{equation}
is (non-decreasing) the cumulative hazard function.

If observations are censored at time $\tau$, then the likelihood of $n$ observations, $n_0$ observed and $n-n_0$ censored, is
\begin{equation}
   f(\{t_i\}_{i=1}^n | \lambda) = \left[\prod_{i=1}^{n_0} \lambda(t_i) \right]
                                  e^{-\sum_{i=1}^{n_0} \Lambda(t_i) - (n-n_0)\Lambda(\tau)}
   (\#eq:model-likef)
\end{equation}

## Simulation

To Draw $n$ observations from the model according to \@ref(eq:model-likef): 

  * Draw $n$ iid values $\{u_i\}_{i=1}^n$ from $\text{Uniform}(0,1)$
  * For each $i=1,\ldots,n$ solve
    \[
        \Lambda(t^\ast_i) = -\log u_i
    \]
  * Set $t_i = \text{min}(t^\ast_i,\tau)$

## Priors on the hazard rate function

Let $G(\cdot)$, $G_1(\cdot)$ and $G_2(\cdot)$ be non-normalised measures with support on the positive real line.  Then consider four cases:

  1. Increasing Failure Rate (IFR) (After @LoWeng:1989)
    \begin{equation}
       \lambda(t|G) = \lambda_0 + \int_0^t G({\rm d}u)
       (\#eq:ifr)
    \end{equation}
  1. Decreasing Failure Rate (DFR) (After @LoWeng:1989)
    \begin{equation}
       \lambda(t|G) = \lambda_0 + \int_t^\infty G({\rm d}u)
       (\#eq:dfr)
    \end{equation}
  1. Lo-Weng Bathtub (LWB) (After @LoWeng:1989)
    \begin{equation}
       \lambda(t|G) = \lambda_0 + \int_0^{|t-a|} G({\rm d}u)
       (\#eq:low-weng-bathtub)
    \end{equation}
    for constant $a>0$.
  1. Superposition Bathtub (SBT) (After @Arnold_Chukova_Hayakawa_APARM:2020)
    \begin{equation}
       \lambda(t|\lambda_0, G_1, G_2) = \lambda_0 + \int_t^\infty G_1({\rm d}u) + \int_0^t G_2({\rm d}u)
       (\#eq:superposition-bathtub)
    \end{equation}
    Here $G_1$ is DFR (determines the early behaviour) and $G_2$ is IFR (determines the late behaviour).
  1. Mixture Bathtub (MBT): Specify through the survival function
    \begin{equation}
       \bar{F}(t|\lambda(\cdot|\pi, G_1, G_2)) = \pi \bar{F}_{\rm DFR}(t|G_1) + (1-\pi) \bar{F}_{\rm IFR}(t|G_2)
       (\#eq:mixture-bathtub)
    \end{equation}
    Where $0<\pi<1$ and $\bar{F}_{\rm DFR}(t|G)$ and $\bar{F}_{\rm IFR}(t|G)$ are the respective survival functions 
    from the IFR and DFR cases defined above. 
  1. Log-convex (LCV)
    \begin{equation}
       \frac{d\log\lambda(t|G)}{dt} = w_0 + \int_0^t G({\rm d}u)
    \end{equation}
    which implies that
    \begin{equation}
       \lambda(t|G) = \lambda_0 e^{w_0t + \int_0^t (t-u) G({\rm d}u)}
       (\#eq:lcfr)
    \end{equation}
    Here the constant $w_0\geq 0$ gives an IFR distribution, but $w_0<0$ gives a bathtub.

Gamma Process Priors can be placed on the mixing measures $G$, $G_1$ and $G_2$ and implemented under the stick-breaking convention.   Specifically if $G|\alpha H_0,\beta \sim \text{GaPP}(\alpha H_0,\beta)$ for probability measure $H_0$ (with support on the positive real line) and postive constants $\alpha$, $\beta$, then a draw from the prior can be written
\begin{eqnarray}
   \theta_k | H_0 &\overset{\rm iid}{\sim}& H_0 \qquad \text{for $k=1,2,\ldots$}\\
   v_k | \alpha &\overset{\rm iid}{\sim}& \text{Be}(1,\alpha)\\
   \tilde{w}_k &=& v_k \prod_{\ell<k} (1-v_\ell)\\
   \gamma &\sim& \text{Ga}(\alpha,\beta)\\
   G({\rm d}u) &=& \gamma\sum_{k=1}^\infty \tilde{w}_k \delta_{\theta_k}({\rm d}u)
   (\#eq:gsum-inf)
\end{eqnarray}
so that
\begin{eqnarray}
  \int_0^t G({\rm d}u) &=& \gamma \sum_{k=1}^\infty \tilde{w}_k I(\theta_k<t)\\
  \int_t^\infty G({\rm d}u) &=& \gamma \sum_{k=1}^\infty \tilde{w}_k I(\theta_k>t)
\end{eqnarray}
and
\begin{eqnarray}
  \int_0^t \int_0^u G({\rm d}v) &=& \gamma \sum_{k=1}^\infty \tilde{w}_k \text{max}(t-\theta_k,0)\\
  \int_0^t \int_u^\infty G({\rm d}v) &=& \gamma \sum_{k=1}^\infty \tilde{w}_k \text{min}(t,\theta_k)
\end{eqnarray}

The weights $\tilde{w}_k$ are stochastically decreasing in $k$, which means that it is possible to truncate the sum in 
eq. \@ref(eq:gsum-inf) after $K$ terms.  In that case we have
\begin{eqnarray}
   \theta_k | H_0 &\overset{\rm iid}{\sim}& H_0 \qquad \text{for $k=1,2,\ldots,K$}\\
   v_k | \alpha &\overset{\rm iid}{\sim}& \text{Be}(1,\alpha) \qquad \text{for $k=1,2,\ldots,K-1$}\\
   \tilde{w}_k &=& v_k \prod_{\ell<k} (1-v_\ell)  \qquad \text{for $k=1,2,\ldots,K-1$}\\
   \tilde{w}_K &=& 1-\sum_{k=1}^{K-1} \tilde{w}_k = \prod_{\ell=1}^{K-1} (1-v_\ell)\\
   \gamma &\sim& \text{Ga}(\alpha,\beta)\\
   G({\rm d}u) &=& \gamma\sum_{k=1}^K \tilde{w}_k \delta_{\theta_k}({\rm d}u)
   (\#eq:gsum-inf)
\end{eqnarray}
A fully hierarchical Bayesian model for $G$ can be completed by choosing a form for $H_0$, with parameters $\phi$, and specifying suitable priors for $\alpha$, $\beta$, $\gamma$ and $\phi$.  For example if $H_0=\text{Exp}(\phi)$ then
\begin{eqnarray}
   \alpha &\sim& \text{Ga}(a_1,a_2)\\
   \beta  &\sim& \text{Ga}(b_1,b_2)\\
   \phi   &\sim& \text{Ga}(f_1,f_2)
\end{eqnarray}

## Specific hazard rate functions

In what follows we write Write $w_k = \gamma\tilde{w}_k$, and also 
define $k^\ast(t) = \text{max}\{k\in (1,\ldots,K)\,:\,\theta^\ast_k\leq t\}$ which implies that
\begin{equation}
    \theta^\ast_{k^\ast(t)} \leq t < \theta^\ast_{k^\ast(t)+1}
\end{equation}
And we may need to re-index $\{(w_k,\theta_k)\}_{k=1}^K$ as $\{(w^\ast_k,\theta_k^\ast)\}_{k=1}^K$ such that
\[
    \theta^\ast_0 \equiv 0 < \theta^\ast_1 < \theta^\ast_2 < \ldots < \theta^\ast_K
\]
(Note that the $\theta_k$ values are almost surely distinct.)  

We also define the following partial sums of the ordered weights and locations:
\begin{eqnarray}
   C^\ast_\ell &=& \sum_{k=1}^\ell w^\ast_k = \gamma - \sum_{k=\ell+1}^K w_k^\ast\\
   D^\ast_\ell &=& \sum_{k=1}^\ell w^\ast_k \theta^\ast_k
\end{eqnarray}

### IFR case

\begin{eqnarray}
   \lambda(t|\lambda_0,\gamma,\mathbf{\theta},\mathbf{v})
      &=& \lambda_0 + \gamma \sum_{k=1}^K \tilde{w}_k I(\theta_k<t)\\
   \Lambda(t|\lambda_0,\gamma,\mathbf{\theta},\mathbf{v})
      &=& \lambda_0 t + \gamma \sum_{k=1}^K \tilde{w}_k \text{max}(0,t-\theta_k)
\end{eqnarray}
If we define $\Lambda^\ast_k=\Lambda(\theta^\ast_k)$ with $\theta^\ast_0=0$ and $\Lambda^\ast_0=0$, then $\Lambda$ is piecewise linear between the points $\{(\theta^\ast_k,\Lambda^\ast_k)\}_{k=0}^K$.   If $U\sim\text{Uniform}(0,1)$ then $T$ is a draw from this distribution if
\begin{eqnarray}
   T = \frac{-\log U + D^\ast_{k^\ast}}{\lambda_0 + C^\ast_{k^\ast}}
\end{eqnarray}
where $k^\ast = \text{max}\{k\in\{1,\ldots,K\}\,:\,\Lambda^\ast_{k}\leq -\log U\}$.

The prior includes the prior for $G\sim\text{GaPP}(\alpha H_0,\beta)$ and a suitable prior for 
$\lambda_0|G$:
\[
   \frac{\lambda_0}{\gamma} \sim \text{Exp}(\nu)
\]

### DFR case

\begin{eqnarray}
   \lambda(t|\lambda_0,\gamma,\mathbf{\theta},\mathbf{v})
      &=& \lambda_0 + \gamma \sum_{k=1}^K \tilde{w}_k I(\theta_k>t)\\
   \Lambda(t|\lambda_0,\gamma,\mathbf{\theta},\mathbf{v})
      &=& \lambda_0 t + \gamma \sum_{k=1}^K \tilde{w}_k \text{min}(t,\theta_k)
\end{eqnarray}
Following the same protocol as in the IFR case: 
if we define $\Lambda^\ast_k=\Lambda(\theta^\ast_k)$ with $\theta^\ast_0=0$ and $\Lambda^\ast_0=0$, then $\Lambda$ is piecewise linear between the points $\{(\theta^\ast_k,\Lambda^\ast_k)\}_{k=0}^K$.   If $U\sim\text{Uniform}(0,1)$ then $T$ is a draw from this distribution if
\begin{eqnarray}
   T = \frac{-\log U - D^\ast_{k^\ast}}{\lambda_0 + \gamma - C^\ast_{k^\ast}}
\end{eqnarray}
where $k^\ast = \text{max}\{k\in\{1,\ldots,K\}\,:\,\Lambda^\ast_{k}\leq -\log U\}$

As in the DFR case, the prior includes the prior for $G\sim\text{GaPP}(\alpha H_0,\beta)$ and a 
suitable prior for $\lambda_0|G$:
\[
   \frac{\lambda_0}{\gamma} \sim \text{Exp}(\nu)
\]

### LWB case

\begin{eqnarray}
   \lambda(t|\lambda_0,a,\gamma,\mathbf{\theta},\mathbf{v})
      &=& \lambda_0 + \gamma \sum_{k=1}^K \tilde{w}_k I(0<\theta_k<|t-a|)\\
      &=& \lambda_0 + 
          \begin{cases}
          \gamma\sum_{k=1}^K \tilde{w}_k I(t<a-\theta_k)\;\; & \text{if $t\leq a$}\\
          \gamma\sum_{k=1}^K \tilde{w}_k I(t>a+\theta_k)\;\; & \text{if $t> a$}
          \end{cases}\\
   \Lambda(t|\lambda_0,a,\gamma,\mathbf{\theta},\mathbf{v})
      &=& \lambda_0 t + 
          \begin{cases}
          \gamma\sum_{k=1}^K \tilde{w}_k I(\theta_k<a)\text{min}(t,a-\theta_k)\;\; 
                          & \text{if $t\leq a$}\\
          \gamma\sum_{k=1}^K \tilde{w}_k 
          \left[\text{max}(0,a-\theta_k)+\text{max}(t-a-\theta_k,0)\right]\;\; & \text{if $t> a$}
          \end{cases}
\end{eqnarray}
Combine the $\{(\theta_{k},w_k)\}_{k=1}^K$ values with one further location, weight pair: $(a,0)$.  Order the locations, and their associated weights, forming the set $\{(w^{\ast\ast}_k,\theta^{\ast\ast}_k)\}_{k=1}^{K+1}$.  Compute the cumulative hazard function values $\Lambda^{\ast\ast}_k=\Lambda(\theta^{\ast\ast}_k)$.  Then if $U\sim\text{Uniform}(0,1)$ then $T$ is a draw from this distribution if
\begin{eqnarray}
   T = \theta^{\ast\ast}_k + \frac{\theta^{\ast\ast}_{k^\ast+1}-\theta^{\ast\ast}_{k^\ast}}{
                                   \Lambda^{\ast\ast}_{k^\ast+1}-\Lambda^{\ast\ast}_{k^\ast}}
                                   \left(-\log U-\Lambda^{\ast\ast}_{k^\ast}\right)
\end{eqnarray}
where $k^\ast = \text{max}\{k\in\{1,\ldots,K_1+K_2\}\,:\,\Lambda^{\ast\ast}_{k}\leq -\log U\}$

Once again the prior includes the prior for $G\sim\text{GaPP}(\alpha H_0,\beta)$ and a suitable prior for $\lambda_0|G$:
\[
   \frac{\lambda_0}{\gamma} \sim \text{Exp}(\nu)
\]


### SBT case

Replicate the prior structure for $G$ for each of $G_1$ and $G_2$, 
indexing the parameters for each component with the label $c=1,2$, 
and add a prior for $\lambda_0|G_2$:
\begin{equation}
    \frac{\lambda_0}{\gamma_2} \sim \text{Exp}(\nu)
\end{equation}

\begin{eqnarray}
   \lambda(t|\lambda_0,\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)
      &=& \lambda_0 + \gamma_1 \sum_{k=1}^{K_1} \tilde{w}_{1k} I(\theta_{1k}>t)
                    + \gamma_2 \sum_{k=1}^{K_2} \tilde{w}_{2k} I(\theta_{2k}<t)\\
   \Lambda(t|\lambda_0,\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)
      &=& \lambda_0 t + \gamma_1 \sum_{k=1}^{K_1} \tilde{w}_{1k} \text{min}(t,\theta_{1k})
                      + \gamma_2 \sum_{k=1}^{K_2} \tilde{w}_{2k} \text{max}(0,t-\theta_{2k})
\end{eqnarray}
Extend the IFR/DFR approach: pool the $\theta_{1k}$ and $\theta_{2k}$ locations and order them, and their associated weights, forming the set $\{(w^{\ast\ast}_k,\theta^{\ast\ast}_k)\}_{k=1}^{K_1+K_2}$.  Compute the cumulative hazard function values $\Lambda^{\ast\ast}_k=\Lambda(\theta^{\ast\ast}_k)$.  Then if $U\sim\text{Uniform}(0,1)$ then $T$ is a draw from this distribution if
\begin{eqnarray}
   T &=& \theta^{\ast\ast}_{k^\ast} 
         + \frac{\theta^{\ast\ast}_{k^\ast+1}-\theta^{\ast\ast}_{k^\ast}}{
                                   \Lambda^{\ast\ast}_{k^\ast+1}-\Lambda^{\ast\ast}_{k^\ast}}
                                   \left(-\log U-\Lambda^{\ast\ast}_{k^\ast}\right)
\end{eqnarray}
where $k^\ast = \text{max}\{k\in\{1,\ldots,K_1+K_2\}\,:\,\Lambda^{\ast\ast}_{k}\leq -\log U\}$

### MBT case

\begin{eqnarray}
   \lambda(t|\pi,
            \lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,
            \lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)
   &=& \frac{f(t|\pi,
                  \lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,
                  \lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)}{
             \bar{F}(t|\pi,
                  \lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,
                  \lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)}\\
   &=& \frac{\pi \lambda_{\rm DFR}(t| \lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1) 
             \bar{F}_{\rm DFR}(t| \lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1)
             +(1-\pi) \lambda_{\rm IFR}(t| \lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2) 
             \bar{F}_{\rm IFR}(t|\lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)
   }{\pi \bar{F}_{\rm DFR}(t|\lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1)
             +(1-\pi) \bar{F}_{\rm IFR}(t|\lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)}\\
   \Lambda(t|\pi,\lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,
                 \lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)
   &=& -\log(\bar{F}(t|\pi,\lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1,
                           \lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2))\\
   &=& -\log\left(\pi \bar{F}_{\rm DFR}(t|\lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1)
            +(1-\pi) \bar{F}_{\rm IFR}(t|\lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)\right)\\
   &=& -\log\left(\pi e^{-\Lambda_{\rm DFR}(t|\lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1)}
             +(1-\pi) 
             e^{-\Lambda_{\rm IFR}(t|\lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)}\right)
\end{eqnarray}
Simulate as a mixture: draw $c\sim\text{Cat}(\{1,2\};(\pi,1-\pi))$ then if $c=1$ draw from the DFR component (parameters $\lambda_{01},\gamma_1,\mathbf{\theta}_1,\mathbf{v}_1)$), and if $c=2$ draw from the IFR component (parameters $\lambda_{02},\gamma_2,\mathbf{\theta}_2,\mathbf{v}_2)$).

The prior includes the priors for 
$G_1\sim\text{GaPP}(\alpha_1 H_{10},\beta_1)$ 
and $G_2\sim\text{GaPP}(\alpha_2 H_{20},\beta_2)$ 
and suitable priors for $\lambda_{10}|G_1$ and $\lambda_{20}|G_2$ are:
\begin{eqnarray}
   \frac{\lambda_{10}}{\gamma_1} &\sim& \text{Exp}(\nu)\\
   \frac{\lambda_{20}}{\gamma_2} &\sim& \text{Exp}(\nu)
\end{eqnarray}

### Log convex case

\begin{eqnarray}
   \frac{d\log\lambda(t|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v})}{dt} 
   &=& w_0 + \sum_{k=1}^K w_k I(0<\theta_k<t)\\
   \log\lambda(t|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v}) &=& \log\lambda_0 + w_0 t 
                                           + \sum_{k=1}^K w_k \text{max}(0,t-\theta_k)\\
   \lambda(t|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v}) &=& \lambda_0 \exp\left(w_0 t + \sum_{k=1}^K w_k \text{max}(0,t-\theta_k)\right)\\
\end{eqnarray}
Then, switching to the re-indexed set $\{(w^\ast_k,\theta^\ast_k)\}_{k=1}^K$
\begin{eqnarray}
   \Lambda(t|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v}) &=& \sum_{\ell=0}^{k^\ast(t)-1} 
                      \int_{\theta^\ast_\ell}^{\theta^\ast_{\ell+1}} 
                            \lambda(u|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v})\,{\rm d}u\\
                  & & + \int_{\theta^\ast_{k^\ast(t)}}^{t} 
                            \lambda(u|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v})\,{\rm d}u\\
                  &=& \sum_{\ell=0}^{k^\ast(t)-1} 
                      \int_{\theta^\ast_\ell}^{\theta^\ast_{\ell+1}} 
                            \lambda_0\exp\left(w_0 u + \sum_{k=1}^K w^\ast_k \text{max}(0,u-\theta^\ast_k)\right)
                            \,{\rm d}u\\
                  & & + \int_{\theta^\ast_{k^\ast(t)}}^{t} 
                            \lambda_0\exp\left(w_0 u + \sum_{k=1}^K w^\ast_k \text{max}(0,u-\theta^\ast_k)\right)
                            \,{\rm d}u\\
                  &=& \sum_{\ell=0}^{k^\ast(t)-1} 
                      \int_{\theta^\ast_\ell}^{\theta^\ast_{\ell+1}} 
                            \lambda_0\exp\left(w_0 u + \sum_{k=1}^\ell w^\ast_k (u-\theta^\ast_k)\right)
                            \,{\rm d}u\\
                  & & + \int_{\theta^\ast_{k^\ast(t)}}^{t} 
                            \lambda_0\exp\left(w_0 u + \sum_{k=1}^{k^\ast(t)} w^\ast_k (u-\theta^\ast_k)\right)
                            \,{\rm d}u\\
                  &=& \sum_{\ell=0}^{k^\ast(t)-1} \lambda_0 e^{-\sum_{k=1}^\ell w^\ast_k\theta^\ast_k}
                      \frac{e^{(w_0+\sum_{k=1}^\ell w^\ast_k)\theta^\ast_{\ell+1}}
                           -e^{(w_0+\sum_{k=1}^\ell w^\ast_k)\theta^\ast_{\ell}}}{
                           w_0 + \sum_{k=1}^\ell w^\ast_k}\\
                  & & + \lambda_0 e^{- \sum_{k=1}^{k^\ast(t)} w^\ast_k\theta^\ast_k}
                      \frac{e^{(w_0+\sum_{k=1}^{k^\ast(t)} w^\ast_k)t}
                           -e^{(w_0+\sum_{k=1}^{k^\ast(t)} w^\ast_k)\theta^\ast_{k^\ast(t)}}}{
                           w_0 + \sum_{k=1}^{k^\ast(t)} w^\ast_k}\\
                  &=& \sum_{\ell=0}^{k^\ast(t)-1} \frac{\lambda_0 e^{- D^\ast_\ell}}{C^\ast_{0\ell}}
                      \left\{e^{C^\ast_{0\ell}\theta^\ast_{\ell+1}}
                            -e^{C^\ast_{0\ell}\theta^\ast_{\ell}}\right\}\\
                  & & + \frac{\lambda_0 e^{- D^\ast_{k^\ast(t)}}}{C^\ast_{0k^\ast(t)}}
                      \left\{e^{C^\ast_{0k^\ast(t)}t}
                            -e^{C^\ast_{0k^\ast(t)}\theta^\ast_{k^\ast(t)}}\right\}
\end{eqnarray}
where
\begin{eqnarray}
   C^\ast_{0\ell} = w_0 + C^\ast_\ell = \sum_{k=0}^\ell w^\ast_k
\end{eqnarray}
Again following the same protocol as in the IFR case
we define $\Lambda^\ast_k=\Lambda(\theta^\ast_k)$ with $\theta^\ast_0=0$ and $\Lambda^\ast_0=0$.  Then in the interval
$\theta^\ast_{k^\ast(t)}\leq t < \theta^\ast_{k^\ast(t)}$ the function $\Lambda$ has the exponential form
\begin{eqnarray}
   \Lambda(t|\lambda_0,w_0,\gamma,\mathbf{\theta},\mathbf{v})
   &=&
   \Lambda^\ast_{k^\ast(t)} + \frac{\lambda_0 e^{- D^\ast_{k^\ast(t)}}}{C^\ast_{0l^\ast}}
                      \left\{e^{C^\ast_{0k^\ast(t)}t}
                           -e^{C^\ast_{0k^\ast(t)}\theta^\ast_{k^\ast(t)}}\right\}
\end{eqnarray}
It follows that 
If $U\sim\text{Uniform}(0,1)$ then $T$ is a draw from this distribution if
\begin{eqnarray}
   T = \frac{1}{C^\ast_{0k^\ast}}\log\left[
         e^{C^\ast_{0k^\ast}\theta^\ast_{k^\ast}}
       + C^\ast_{0k^\ast}\frac{-\log U - \Lambda^\ast_{k^\ast}}{\lambda_0 e^{-D^\ast_{k^\ast}}}
       \right]
\end{eqnarray}
where $k^\ast = \text{max}\{k\in\{1,\ldots,K\}\,:\,\Lambda^\ast_{k}\leq -\log U\}$

Priors are needed for $\lambda_0$ and $w_0$.  Suitable priors are
\begin{eqnarray}
   \frac{1}{\gamma}\log\lambda_0 &\sim& \text{Normal}(0,\nu^{-2})\\
   \frac{w_0}{\gamma} &\sim& \text{Normal}(0,\nu^{-2}) 
\end{eqnarray}

## Samplers

### IFR case

  * $\theta_k$
  
  Full conditional:
  \begin{eqnarray}
  f(\theta_k | - ) &\propto& 
            \left[\prod_{i=1}^{n_0} \lambda(t_i | \mathbf{\theta}, \mathbf{v})\right]
            \exp\left[-\sum_{i=1}^n\Lambda(t_i | \mathbf{\theta}, \mathbf{v})\right]\\
            &&\times \exp(-\phi\theta_k)
  \end{eqnarray}
  Proposal: Select $k\sim\text{Cat}(\tilde{\mathbf{w}}(\mathbf{v}))$ and then propose 
  $\theta^\ast_k$ by a log Normal random walk Metropolis step: 
  $\log\theta_k^\ast | \log\theta_k, \sigma_\theta^2 \sim N(\log\theta_k, \sigma_\theta^2)$
  \begin{eqnarray}
  q(k, \theta^\ast_k | \theta_k, \mathbf{v}) 
                    &\propto& \frac{\tilde{w}_k}{\theta^\ast_k} 
                    e^{-(\log\theta^\ast_k-\log\theta_k)^2/(2\sigma_\theta^2)}
  \end{eqnarray}
  Proposal ratio:
  \begin{eqnarray}
  r &=& \frac{f(\theta^\ast_k|-)}{f(\theta_k|-)} 
        \frac{q(k, \theta_k | \theta^\ast_k, \mathbf{v})}{q(k, \theta^\ast_k | \theta_k, \mathbf{v})}\\
    &=& \left[\prod_{i=1}^{n_0} \frac{\lambda(t_i | \mathbf{\theta}^\ast, \mathbf{v})}{
                                      \lambda(t_i | \mathbf{\theta}, \mathbf{v})}\right]\\
    &&\times \exp\left[-\sum_{i=1}^n\left\{
                        \Lambda(t_i | \mathbf{\theta}^\ast, \mathbf{v})
                       -\Lambda(t_i | \mathbf{\theta}, \mathbf{v})
                        \right\}\right]\\
    &&\times \exp(-\phi(\theta^\ast_k-\theta_k))\\
    &&\times \frac{\theta^\ast_k}{\theta_k}\\
  \log r &=& \left[\sum_{i=1}^{n_0} \{\log\lambda(t_i | \mathbf{\theta}^\ast, \mathbf{v})
                                     -\log\lambda(t_i | \mathbf{\theta}, \mathbf{v})\}\right]\\
    && -\sum_{i=1}^n\left\{
                        \Lambda(t_i | \mathbf{\theta}^\ast, \mathbf{v})
                       -\Lambda(t_i | \mathbf{\theta}, \mathbf{v})
                        \right\}\\
    && -\phi(\theta^\ast_k-\theta_k) + \log\theta^\ast_k-\log\theta_k
  \end{eqnarray}
  
  * $v_k$
  
  Full conditional:
  \begin{eqnarray}
  f(v_k | - ) &\propto& 
            \left[\prod_{i=1}^{n_0} \lambda(t_i | \mathbf{\theta}, \mathbf{v})\right]
            \exp\left[-\sum_{i=1}^n\Lambda(t_i | \mathbf{\theta}, \mathbf{v})\right]\\
            &&\times (1-v_k)^{\alpha-1}
  \end{eqnarray}
  Proposal: Select $k\sim\text{Cat}(\tilde{\mathbf{w}}(\mathbf{v}))$ and then propose 
  $v^\ast_k$ by a logistic Normal random walk Metropolis step: 
  $\text{logit}v_k^\ast | \text{logit}\,v_k, \sigma_v^2 \sim N(\text{logit}\,v_k, \sigma_v^2)$
  \begin{eqnarray}
  q(k, v^\ast_k | v_k, \mathbf{v}_{-k}) 
                    &\propto& \frac{\tilde{w}_k(\mathbf{v})}{v^\ast_k(1-v^\ast_k)} 
                    e^{-(\text{logit}\,v^\ast_k-\text{logit}\,v_k)^2/(2\sigma_v^2)}\\
                    &\propto& \frac{v_k}{v^\ast_k(1-v^\ast_k)} 
                    e^{-(\text{logit}\,v^\ast_k-\text{logit}\,v_k)^2/(2\sigma_v^2)}\\
  \end{eqnarray}
  Proposal ratio:
  \begin{eqnarray}
  r &=& \frac{f(v^\ast_k|-)}{f(v_k|-)} 
        \frac{q(k, v_k | v^\ast_k, \mathbf{v}_{-k})}{q(k, v^\ast_k | v_k, \mathbf{v}_{-k})}\\
    &=& \left[\prod_{i=1}^{n_0} \frac{\lambda(t_i | \mathbf{\theta}, \mathbf{v}^\ast)}{
                                      \lambda(t_i | \mathbf{\theta}, \mathbf{v})}\right]\\
    &&\times \exp\left[-\sum_{i=1}^n\left\{
                        \Lambda(t_i | \mathbf{\theta}, \mathbf{v}^\ast)
                       -\Lambda(t_i | \mathbf{\theta}, \mathbf{v})
                        \right\}\right]\\
    &&\times \left(\frac{v_k^\ast}{v_k}\right)^2\left(\frac{1-v_k^\ast}{1-v_k}\right)^{\alpha}\\
  \log r &=& \left[\sum_{i=1}^{n_0} \{\log\lambda(t_i | \mathbf{\theta}^\ast, \mathbf{v})
                                     -\log\lambda(t_i | \mathbf{\theta}, \mathbf{v})\}\right]\\
    && -\sum_{i=1}^n\left\{
                        \Lambda(t_i | \mathbf{\theta}^\ast, \mathbf{v})
                       -\Lambda(t_i | \mathbf{\theta}, \mathbf{v})
                        \right\}\\
    && +\alpha\log\left(\frac{1-v_k^\ast}{1-v_k}\right)+2\log\left(\frac{v_k^\ast}{v_k}\right)
  \end{eqnarray}
  
  * $\gamma$
  
  Full conditional:
  \begin{eqnarray}
  f(\gamma | - ) &\propto& 
            \left[\prod_{i=1}^{n_0} \lambda(t_i | \mathbf{\theta}, \mathbf{v})\right]
            \exp\left[-\sum_{i=1}^n\Lambda(t_i | \mathbf{\theta}, \mathbf{v})\right]\\
            &&\times ...
  \end{eqnarray}
  
  * $\alpha$
  * $\beta$
  * $\phi$

\begin{equation}
\end{equation}

# Bibliography